% ====================================================================
%+
% SECTION:
%    SolarSystem_OrbitalAccuracy.tex
%
% CHAPTER:
%    solarsystem.tex
%
% ELEVATOR PITCH:
%    How secure is the orbit - is it going to hit us?
%    Libration amplitude distribution for TNOs?
%    Can we find it after X years for further study?
%    Can we identify the source region for NEOs within the main belt?
%-
% ====================================================================

\section{Orbital Accuracy}
\def\secname{\chpname:orbits}\label{sec:\secname}

After initial discovery,

A short preamble goes here. What's the context for this science
project? Where does it fit in the big picture?

How secure is the orbit - is it going to hit us?
Libration amplitude distribution for TNOs?
Can we find it after X years for further study?
Can we identify the source region for NEOs within the main belt?

% --------------------------------------------------------------------

\subsection{Target measurements and discoveries}
\label{sec:\secname:targets}

Describe the discoveries and measurements you want to make.

Now, describe their response to the observing strategy. Qualitatively,
how will the science project be affected by the observing schedule and
conditions? In broad terms, how would we expect the observing strategy



% --------------------------------------------------------------------

\subsection{Metrics}
\label{sec:\secname:metrics}

Quantifying the response via MAF metrics: definition of the metrics,
and any derived overall figure of merit.


% --------------------------------------------------------------------

\subsection{OpSim Analysis}
\label{sec:\secname:analysis}

OpSim analysis: how good would the default observing strategy be, at
the time of writing for this science project?


% --------------------------------------------------------------------

\subsection{Discussion}
\label{sec:\secname:discussion}

Discussion: what risks have been identified? What suggestions could be
made to improve this science project's figure of merit, and mitigate
the identified risks?

\navigationbar
