% ====================================================================
%+
% SECTION:
%    section-name.tex  % eg lenstimedelays.tex
%
% CHAPTER:
%    chapter.tex  % eg cosmology.tex
%
% ELEVATOR PITCH:
%    Explain in a few sentences what the relevant discovery or
%    measurement is going to be discussed, and what will be important
%    about it. This is for the browsing reader to get a quick feel
%    for what this section is about.
%
% COMMENTS:
%
%
% BUGS:
%
%
% AUTHORS:
%    Phil Marshall (@drphilmarshall)  - put your name and GitHub username here!
%-
% ====================================================================

\section{AGN Selection and Census}
\def\secname{\chpname:census}\label{sec:\secname}

\credit{ohadshemmer},
\credit{nielbrandt},
\credit{GordonRichards},
\credit{ScottAnderson},
{\it and others to follow}

% This individual section will need to describe the particular
% discoveries and measurements that are being targeted in this section's
% science case. It will be helpful to think of a ``science case" as a
% ``science project" that the authors {\it actually plan to do}. Then,
% the sections can follow the tried and tested format of an observing
% proposal: a brief description of the investigation, with references,
% followed by a technical feasibility piece. This latter part will need
% to be quantified using the MAF framework, via a set of metrics that
% need to be computed for any given observing strategy to quantify its
% impact on the described science case. Ideally, these metrics would be
% combined in a well-motivated figure of merit. The section can conclude
% with a discussion of any risks that have been identified, and how
% these could be mitigated.

% A short preamble goes here. What's the context for this science
% project? Where does it fit in the big picture?

%\new{From Phil: Large samples of LSST AGN will provide high precision constraints
%on  galaxy evolution models. The science yield will be maximized  by
%increasing the sample size; the accuracy with which the selection
%function is known may also be important. }

The primary goal for AGN science is to maximize the discovery of AGN
with the LSST and construct the largest possible inventory of sources
spanning the widest possible ranges in the redshift-luminosity parameter
space. This, in turn, will provide high-precision constraints on galaxy
evolution models as well as on various cosmological science cases,
such as quasar clustering, the highest redshift quasars, and
strong gravitational lensing.

% --------------------------------------------------------------------

\subsection{Target measurements and discoveries}
\label{sec:\secname:targets}

% Describe the discoveries and measurements you want to make.

% Now, describe their response to the observing strategy. Qualitatively,
% how will the science project be affected by the observing schedule and
% conditions? In broad terms, how would we expect the observing strategy
% to be optimized for this science?

%\new{From Phil: Our target measurements will be inferences of galaxy evolution
%model  parameters, from a joint analysis with the host galaxy
%properties. As an initial proxy, we focus on the quasar luminosity
%function -- and  as a first step towards that, the number counts.}

About $10^7 - 10^8$ AGNs will be selected in the main LSST survey
using a combination of criteria, split broadly into four categories:
colors, astrometry, variability, and multiwavelength matching with
other surveys. The LSST observing strategy will affect mostly the
first three of these categories as described further below. The most
critical measurement is the a reliable and precise redshift for each
source, obtained both from a photometric and an astrometric redshift.

{\bf Colors:} The LSST observing strategy will determine the depth in
each band, as a function of position on the sky, and will thus affect
the color selection of AGNs. This will eventually determine the AGN
$L-z$ distribution and, in particular, may affect the identification
of quasars at $z\gtsim 6$ if, for example, $Y$-band exposures will
not be sufficiently deep.

{\bf Variability:} AGNs can be effectively distinguished from (variable)
stars, and from quiescent galaxies, by exhibiting certain characteristic
variability patterns (e.g., \citet{ButlerandBloom2011}). Non-uniform
sampling may ``contaminate'' the variability signal of AGN candidates.

{\bf Astrometry:} AGNs will be selected among sources having zero
proper motion, within the uncertainties. The LSST cadence may affect
the level of this uncertainty in each band, and may therefore affect
the ability to identify (mostly fainter) AGN.
%
Differential chromatic refraction (DCR), making use of the astrometric
offset a source with emission lines has with respect to a source with
a featureless power-law spectrum, can help in the selection of AGNs
and in confirming their photometric redshifts
\citep{KaczmarczikEtal2009}. The DCR effect is more pronounced at
higher airmasses. Therefore, it could be advantageous to have at least one
visit, per source, at airmass greater than about 1.4. AGN selection
and photometric redshift confirmation may be affected since the LSST
cadence will affect the airmass distribution, in each band, for each
AGN candidate.

% Ideas for Metrics:
% detection - how many can LSST detect based on the luminosity function
% (depends on the depth in each band for single epoch and coadd)
% (how will this change with each DR)? @ohadshemmer

% classification - How many of these will we actually classify as quasars?
% non-simultaneous colors. variability of QSOs (how does depend on
% cadence/baseline/seasonal gaps?)

% --------------------------------------------------------------------

\subsection{Metrics}
\label{sec:\secname:metrics}

Quantifying the response via MAF metrics: definition of the metrics,
and any derived overall figure of merit.

Need to compute the mean $Y$ band magnitude
across the sky for the nominal OpSim. Compare this magnitude to the
one required for identifying $\geq1000$ quasars at $z\geq6$.
Currently, for enigma\_1189, the single-epoch 5-sigma depth in $Y$
is 22.36 mag, and for the final co-added 5-sigma depth is $Y=24.4$.
These limits are deeper than the original predictions.
% (See the AAS poster from 2013: http://www.lsst.org/sites/default/files/221-RC-247.10-AAS_shemmer.pptx.pdf).

% --------------------------------------------------------------------

\subsection{OpSim Analysis}
\label{sec:\secname:analysis}

OpSim analysis: how good would the default observing strategy be, at
the time of writing for this science project?


% --------------------------------------------------------------------

\subsection{Discussion}
\label{sec:\secname:discussion}

Discussion: what risks have been identified? What suggestions could be
made to improve this science project's figure of merit, and mitigate
the identified risks?


% ====================================================================

\navigationbar
