% ====================================================================
%+
% SECTION:
%    MW_Plane.tex
%
% CHAPTER:
%    .tex  % eg cosmology.tex
%
% ELEVATOR PITCH:
%    
% The Plane should be part of DWF
% It's possible some of this material should be moved to other sections, but it's here for now.
% 
% COMMENTS:
%
% AUTHOR:
%    Jay Strader (@caprastro)
%    Chris Britt
% ====================================================================

\section{The Galactic Plane Should Be Part of the Main Deep-Wide-Fast Survey}
\def\secname{MW_Plane}\label{sec:\secname} % For example, replace "keyword" with "lenstimedelays"

\noindent{\it Jay Strader, Chris Britt} % (Writing team)


The current baseline cadence (${\tt enigma\_1189}$) partially excludes the Galactic Plane from the deep-wide-fast survey and instead adopts a nominal 30 visits 
per filter as part of a special proposal. The basic justification for this cadence is that the crowding and extinction in the plane makes it impossible to 
achieve many of the extragalactic science goals of the survey. However, much Milky Way science can \emph{only} be accomplished in the plane, so it is not 
prima facie obvious that this area should be excluded from the deep-wide-fast survey. Careful consideration of science goals and metrics is necessary to 
determine whether the plane should be addressed in a special proposal---and the details of that proposal. Here we briefly highlight the science cases that 
focus on the Plane and show that despite the challenges, there is compelling support for including the Galactic Plane in the deep-wide-fast survey.

The science cases emphasized below do not require independent single epoch or combined photometry down to the nominal survey magnitude limits. However, it is 
still desirable to determine quantitative measurements of the effects of crowding and extinction on photometry, proper motions, and parallax as a function of 
Galactic latitude and filter rather than rough qualitative estimates. This will require image simulations through ImSim at a range of representative Galactic 
latitudes that are processed through the stack.

\subsection{Unique Galactic Transients and Variables}

Given the challenges of observing in the Plane, a simple increase of the survey area is insufficient justification for standard deep-wide-fast coverage. 
However, what is relevant is that many classes of important transients and variable stars are mostly or exclusively found in this region. Here we describe a 
representative (but non-exhaustive) list of these variables and transients.


\subsubsection{Stellar-Mass Black Holes}

Of the millions of stellar-mass black holes formed through the collapse of massive stars over the lifetime of the Milky Way, only $\sim 20$ have been 
dynamically confirmed through spectroscopic measurements
\citep[e.g.,][]{2015arXiv151008869C}.  Many questions central to modern
astrophysics can only be answered by enlarging this sample: which stars produce
neutron stars and which black holes; whether there is a true gap in mass
between neutron stars and black holes; whether supernova explosions result in
large black hole kicks.

All of these known stellar-mass black holes were discovered  as bright X-ray sources, either due to persistent emission or as transient ``X-ray novae" 
due to an increase in the mass accretion rate within the standard  disk instability model. This introduces a strong bias in the known black hole candidates:
only those with short outburst recurrence times and whose peak X-ray luminosity reaches $\sim 10^{35}$--$10^{36}$ erg/s can be detected. 
Short-period binaries should be preferentially faint
\citep{2014MNRAS.437.3087K} and may be mostly or entirely missed in current
samples, as evidenced by the current orbital period distribution of black hole
binaries. While the historical rate of these detected outbursts is $\sim 1$ per
year, the number of systems detected by LSST may be much higher since the
majority of systems should be short period and X-ray faint, and hence
undetected by X-ray all sky monitors. Careful simulations are required to
robustly estimate the rate of new black hole transients which LSST could
detect.  These outbursts are optically bright, from $4-8$ magnitudes reaching
$M_{V}\sim 0$, and so should be easily visible at large distances even near the
Plane. Some of the optical light is contributed by the jet, which is quenched
at the peak of the outburst, returning later on as the source fades
\citep{2001ApJ...554L.181J}.

More importantly, there is expected to be a large population of black hole
binaries in quiescence with low X-ray luminosities from $\sim
10^{30}$--$10^{33}$ erg/s.  Such systems can be identified as optical variables
that show unique, double-humped ellipsoidal variations of typical amplitude
$\sim 0.2$ mag due to the tidal deformation of the secondary star, which can be
a giant or main sequence star. In some cases analysis of the light curve alone
can point to a high mass ratio between the components, suggesting a black hole
primary; in other cases the accretion disk will make a large contribution to
the optical light which results in intrinsic, random, and fast variations in
the light curve. The disk contribution to optical light can change over time,
and several years of data is necessary to properly subtract the accretion disk
contribution in order to properly fit ellipsoidal veriations
\citep{2010ApJ...710.1127C}.  These variations could provide an additional way
to distinguish them from contact binaries, and simulations to determine what
sampling is necessary to distinguish them in this way is needed. Additional
follow up on objects discovered by LSST would be required to robustly identify
the nature of such an accreting binary. Some of the metrics used in \S 6.5 will
be of use here as well. 

Most black hole candidates have been identified toward the center of the Galaxy, in the Plane: 68\% of black hole candidates are located within $5^{\circ}$ of the Plane and 92\% within $10^{\circ}$ of the Plane. Therefore, including the Plane in the deep-wide-fast survey offers the best route to identifying the hundreds of quiescent black hole binaries that should be present. The brighter sources will be amenable to spectroscopy with the current generation of 4-m to 10-m telescopes to dynamically confirm new black holes; spectroscopy of all candidates should be possible with the forthcoming generation of large telescopes.

\subsubsection{Other Compact Binaries}

The methods discussed above for black holes can also be used to identify white
dwarf and especially neutron star binaries, which again are preferentially
found in the dense regions of the Galaxy.  In some cases the secondaries can
show clear ellipsoidal variations and be identified and studied as periodic
variables; in other cases the accretion disk light can dominate, resulting in
quasi-periodic or non-periodic light curves. Nonetheless, cross-matching LSST
variables against the all-sky eROSITA catalog (reaching typical X-ray fluxes of
$10^{31}$--$10^{32}$ erg/s at 8 kpc) will reveal hundreds of new neutron star
X-ray binaries \citep{2012arXiv1209.3114M}.

From studies of \emph{Fermi} $\gamma$-ray sources, it is clear that pulsar
surveys miss a huge fraction of pulsars in close binary systems (probably due
to the effects of enshrouding material) but that targeted follow-up can detect
such pulsars \citep{2012arXiv1205.3089R}. Combining these observations with
optical spectroscopic data gives the equivalent of a double-lined spectroscopic
binary, and can often yield accurate masses for both the neutron star and the
secondary. Further, gold-standard Shapiro-delay masses for neutron stars
require edge-on orientations that only arise from large samples of objects.
Such studies offer the best chance to probe the upper end of the neutron star
mass distribution; the maximum mass of neutron stars determines the equation of
state of dense matter and is of central interest to both astronomers and
nuclear physicists. Additionally, proper motions of neutron star X-ray binaries
should be easily detectable; dynamical kicks from supernovae, though poorly
understood, are on the order of hundreds km\,s$^{-1}$. It is unclear what, if
any, dynamical kicks black holes may undergo at birth. A thorough understanding
of the true distribution of dynamical kicks from supernovae will place strong
observational constraints on the internal physics of Type II supernovae. 

A great many dwarf novae from white dwarf binaries will be discovered routinely
by LSST, preferentially towards low Galactic latitudes. Dwarf novae can be
standardizable candles \citep{2011MNRAS.411.2695P} though they tend to be fainter
than stars on the instability strip, even in outburst ($M_{V}\sim 3-6$). The
total number of dwarf novae is, however, poorly understood---theoretical
estimates routinely yield a significantly higher number than are observed in
the solar neighborhood. The selection biases and low number density make this
comparison difficult, but understanding the true specific frequency of these
systems provides a key check on common envelope evolution, which is poorly
understood and has a large impact on, for example, LIGO event rates. LSST will
detect dwarf novae, which last at least several days, out to kpc scales. This
will allow a test of not only the number of cataclysmic variables, but also of
the 3D distribution within the Galaxy and dependence on metallicity gradients
\citep{2015MNRAS.448.3455B}. The cadence of observations is critical in obtaining an
accurate measure of the population of cataclysmic variables, as a long baseline
is necessary to recover low duty cycle systems while observations spaced too
far apart risk leaving windows large enough to allow short outbursts to go
undetected. Appropriate simulations of what science returns is allowed by
various cadence schedules are critical.

\subsubsection{A Milky Way Supernova}

A supernova in the Milky Way would be among the most important astronomical
events of our lifetime, with enormous impacts on stellar astrophysics, compact
objects, nucleosynthesis, and neutrino and gravitational wave astronomy. The
estimated rate of supernovae (both core-collapse and Type Ia) in the Milky Way
is about 1 per 20--25 years \citep{2013ApJ...778..164A}; hence there is a
40--50\% chance that this would occur during the 10-year LSST survey.

If fortunate, such an event will be located relatively close to the Sun and
will be an easily observed (perhaps even naked-eye) event. However, we must be
cognizant of the likelyhood that the supernova could go off in the mid-Plane
close to the Galactic Center or on the other side of the Milky Way---both
regions covered by LSST. While any core-collapse event will produce a
substantial neutrino flux, alerting us to its existence, such observations will
not offer precise spatial localization. The models of
\citet{2013ApJ...778..164A} indicate that LSST is the \emph{only} planned
facility that can offer an optical transient alert of nearly all Galactic
supernovae. Even if the supernova is not too faint, LSST will likely be the
sole facility with synoptic observations preceding the explosion, providing
essential photometric data leading up to the event---but only if LSST covers
the Plane at a frequent cadence.


\subsubsection{Novae}

Only $\sim 15$ novae (explosions on the surfaces of white dwarfs) are
discovered in the Milky Way each year, while observations of external galaxies
show that the rate should be a factor of $\sim 3$ higher
\citep{2014ASPC..490...77S}.  Evidently, we are missing 50--75\% of novae due
to their location in crowded, extinguished regions, where they are not bright
enough to be discovered at the magnitude limits of existing transient surveys.
Fundamental facts about novae are unknown: how much mass is ejected in typical
explosions; whether white dwarfs undergoing novae typically gain or lose mass;
whether the binary companion is important in shaping the observed properties of
nova explosions.  Novae can serve as scaled-down models of supernova explosions
that can be tested in detail, e.g., in the interaction of the explosion with
circumstellar material \citep[e.g.,][]{2015arXiv151007662C}.  Further, since
accreting white dwarfs are prime candidates as progenitors of Type Ia
supernovae, only detailed study of novae can reveal whether particular systems
are increasing toward the Chandrasekhar mass as necessary in this scenario.

Most novae occur in the Galactic Plane and Bulge, and therefore the quick
identification of novae is best enabled by the standard deep-wide-fast cadence.
These events will trigger multi-wavelength follow-up ranging from the radio to
X-ray and $\gamma$-rays; these data are necessary for accurate measurements of
the ejected mass.

\subsection{Microlensing}

Gould (2013) shows that coverage of the Plane with the standard deep-wide-fast
cadence would be, in essence, an effective intra-disk microlensing survey (in
which disk stars are lensed by other objects in the disk, such as exoplanets,
brown dwarfs, or compact objects). The lower stellar density compared to past
bulge-focused microlensing surveys would be offset by the larger area covered
by LSST. The predicted rate of high magnification microlensing events that are
very sensitive to planets would be $\sim 25$ per year. This survey would be
able to detect planets at moderate distances from their host stars, a regime
poorly probed by standard Doppler and transit techniques. The LSST data alone
would not be sufficient: the detection of a slow ($\sim$ days) timescale
increase in brightness of a disk star would need to trigger intensive
photometric observations from small (1-m to 2-m class) telescopes that would
observe at high cadence for the 1--2 months of the microlensing event. This
science \emph{requires} the standard deep-wide-fast cadence (with typical
inter-night visit times of a few days) to catch lensing events as they start to
brighten.


\subsection{A Three-Dimensional Dust Map of the Milky Way}

The Pan-STARRS1 survey (PS1) has produced a three-dimensional dust map of the
region of the sky covered in their 3$\pi$ survey (which excludes a large part
of the Galactic Plane toward the south). Such maps are necessary to accurately
measure the intrinsic luminosities and colors of both Galactic and
extragalactic sources. The PS1 map \citep{2011MNRAS.411.2695P} saturates at
extinctions $E(B-V) > 1.5$ as their tracer stars fall out of the survey
catalogs fainter than $g\sim 22$, meaning that this high-fidelity map does not
extend uniformly to within a few degrees of the midplane. In addition, it only
extends to a distance of about 4.5 kpc. Deep LSST data will allow this map to
be extended to much higher extinctions and larger distances. Owing to the high
extinction and the use of blue filters, this project is less affected by
crowding than other projects requiring photometry in the Plane. Nonetheless,
quantiative estimates of the expected photometric accuracy in coadded $u$ and
$g$ images at low Galactic latitude are desirable.


% --------------------------------------------------------------------

% \subsection{Target measurements and discoveries}
% \label{sec:\secname:targets}

% Describe the discoveries and measurements you want to make.

% Now, describe their response to the observing strategy. Qualitatively,
% how will the science project be affected by the observing schedule and
% conditions? In broad terms, how would we expect the observing strategy
% to be optimized for this science?


% --------------------------------------------------------------------
%
 \subsection{Metrics}
 \label{sec:\secname:metrics}
%

Using proposed variable star metrics (\S 6.3), the recovery of accurate periods for variables discussed above should be compared between the baseline cadence and the
results of the ``Normal Plane" OpSim proposal (in which the portion of the Plane falling within the deep-wide-fast declination limits is observed at the
standard deep-wide-fast cadence).

Several focused surveys on the Bulge with the wide-field MOSAIC II and DECam instruments have demonstrated that \emph{differential} photometry is possible in
these very crowded regions down to close to the LSST single epoch photometry
limits in $r$ and $i$ \citep[e.g.,][]{2014ApJS..214...10B}.
Therefore, periods and amplitudes for variables can be accurately measured even
if precise absolute photometry is not available for every star (even though
such photometry should be possible down to at least the main sequence turnoff;
Rich 2015). In the Plane, away from the Bulge, the crowding is less extreme,
and quantitative limits should be calculated for photometry, proper motion, and
parallax measurements through image simulations as noted above.

%
%
% --------------------------------------------------------------------

%\subsection{OpSim Analysis}
%\label{sec:\secname:analysis}


% --------------------------------------------------------------------

% \subsection{Discussion}
% \label{sec:\secname:discussion}
%
% Discussion: what risks have been identified? What suggestions could be
% made to improve this science project's figure of merit, and mitigate
% the identified risks?
%
%
% ====================================================================

\navigationbar
