% ====================================================================
%+
% SECTION:
%    section-name.tex  % eg lenstimedelays.tex
%
% CHAPTER:
%    chapter.tex  % eg cosmology.tex
%
% ELEVATOR PITCH:
%    Explain in a few sentences what the relevant discovery or
%    measurement is going to be discussed, and what will be important
%    about it. This is for the browsing reader to get a quick feel
%    for what this section is about.
%
% COMMENTS:
%
%
% BUGS:
%
%
% AUTHORS:
%    Phil Marshall (@drphilmarshall)  - put your name and GitHub username here!
%-
% ====================================================================

\section{The spatial structure of the Milky Way}
\def\secname{MW_spatial_structure}\label{sec:\secname} % For example, replace "keyword" with "lenstimedelays"

\noindent{\it Author Name(s)} % (Writing team)

% This individual section will need to describe the particular
% discoveries and measurements that are being targeted in this section's
% science case. It will be helpful to think of a ``science case" as a
% ``science project" that the authors {\it actually plan to do}. Then,
% the sections can follow the tried and tested format of an observing
% proposal: a brief description of the investigation, with references,
% followed by a technical feasibility piece. This latter part will need
% to be quantified using the MAF framework, via a set of metrics that
% need to be computed for any given observing strategy to quantify its
% impact on the described science case. Ideally, these metrics would be
% combined in a well-motivated figure of merit. The section can conclude
% with a discussion of any risks that have been identified, and how
% these could be mitigated.

A short preamble goes here. What's the context for this science
project? Where does it fit in the big picture?

% --------------------------------------------------------------------

\subsection{Target measurements and discoveries}
\label{sec:\secname:targets}

Describe the discoveries and measurements you want to make.

Now, describe their response to the observing strategy. Qualitatively,
how will the science project be affected by the observing schedule and
conditions? In broad terms, how would we expect the observing strategy
to be optimized for this science?


% --------------------------------------------------------------------

\subsection{Metrics}
\label{sec:\secname:metrics}

Quantifying the response via MAF metrics: definition of the metrics,
and any derived overall figure of merit.

Highly incomplete set of examples:

{\bf Metric 1: Fraction of outer disk population ($\gtrsim 15$~kpc) correctly identified}
\begin{itemize}
  \item Scaling metallicity estimate error from program requirements per exposure, photometric precision per filter for main sequence objects
    \item Age precision estimates from photometry and errors in these estimates ($\sim 1$~Gy);
      \item Photometric distance error estimates given apparent magnitude, colors and observing strategy ($\sim 1$~kpc)
        \item Synthetic stellar population including age and metallicity distribution as a function of galactocentric radius;
          \item Monte Carlo simulation of recovery of population
            characteristics given input survey strategy
\end{itemize}

{\bf Metric 2: Accuracy of velocity-metallicity relation for disk stars.} (e.g. Loebman et al. 2011). Dependencies:
\begin{itemize}
  \item Scaling for metallicity and age estimates from program requirements for single exposure pairs;
    \item Proper motion precision scaling estimates (crowded fields?); 
  \item Analytic models for velocity-metallicity relations for synthetic population
\end{itemize} 

{\bf Metric 3: Accuracy of shape parameters for the thick disk population:} Dependencies:
\begin{itemize}
  \item Photometric errors scaled from single-exposure-pair;
    \item Analytic prescriptions for multi-population Milky Way (thin-disk, thick-disk, bulge, halo)
      \item Monte carlo simulation of recovered parameters for a representative sample of thick-disk structural parameters (including no thick disk; e.g. Bovy et al.); ability to sample the same objects in each simulation with the different OpSim run outputs. 
\end{itemize}

% --------------------------------------------------------------------

\subsection{OpSim Analysis}
\label{sec:\secname:analysis}

OpSim analysis: how good would the default observing strategy be, at
the time of writing for this science project?


% --------------------------------------------------------------------

\subsection{Discussion}
\label{sec:\secname:discussion}

Discussion: what risks have been identified? What suggestions could be
made to improve this science project's figure of merit, and mitigate
the identified risks?


% ====================================================================

\navigationbar
