% --------------------------------------------------------------------

\chapter[Cosmology]{Accurate Cosmological Measurements on the Largest Scales}
\def\chpname{cosmo}\label{chp:\chpname}

Chapter editors:
\credit{egawiser},
\credit{MichelleLochner}.

Contributing authors:
\credit{HumnaAwan},
\credit{egawiser},
\credit{pkurczynski},
\credit{rhiannonlynne},
\credit{jmeyers314},
\credit{tonytyson},
\credit{MelissaGraham},
\credit{SamSchmidt},
\credit{connolly},
\credit{ivezic},
\credit{jhrlsst},
\credit{rbiswas4},
\credit{sethdigel},
\credit{astrofoley},
\credit{lgalbany},
\credit{pgris},
\credit{ReneeHlozek},
\credit{saurabhwjha},
\credit{RickKessler},
\credit{AlexGKim},
\credit{oferlahav},
\credit{aimalz},
\credit{jasonmcewen},
\credit{janewman-pitt-edu},
\credit{hiranyapeiris},
\credit{kponder},
\credit{rlschuhmann},
\credit{astrostubbs},
\credit{wmwv},
\credit{drphilmarshall},
\credit{tanguita}

% \section*{Summary}
% \addcontentsline{toc}{section}{~~~~~~~~~Summary}
%
% Executive summary goes here, highlighting the primary conclusions from
% the chapter's science cases. This should be abstract length, no more:
% say, 200 words.

% --------------------------------------------------------------------

\section{Introduction}
\label{sec:\chpname:intro}

Cosmology is one of the key science themes for which LSST was
designed. Our goal is to measure cosmological parameters, such as the
equation of state of dark energy or departures from General
Relativity, with sufficient accuracy to distinguish one model from
another and hence drive our theoretical understanding of how the
universe works as a whole. To do this will necessarily involve a
variety of different measurements that can act as cross-checks and break
any parameter degeneracies.

The  Dark Energy Science Collaboration (DESC) has identified five
different cosmological probes enabled by the LSST: weak lensing (WL),
large-scale structure (LSS), Type Ia supernovae (SN), strong lensing
(SL), and clusters of galaxies (CL). In all these cases, the primary concern
is the residual systematic error: the shapes and photometric redshifts of
galaxies and the properties of supernova and lensed quasar light
curves will all need to be measured with extraordinary accuracy in
order to properly harness the statistical power available through LSST. This
accuracy will come from the abundance and heterogeneity of the
individual measurements and the degree to which they can be
modeled and understood. This latter point implies a need for uniformity
in the survey, which enables powerful simplifying assumptions to be made
when calibrating on the largest, cosmologically most important scales.
The need for heterogeneity in the measurements also requires uniformity in
the sense that the nuisance parameters that describe the systematic effects
need to be sampled over as wide a range as possible (e.g., the need to
sample a wide range of roll angles to minimize shape error;
observing conditions to understand photometric errors due to the
changing atmosphere).

In this chapter we look at some of the key measurements planned by DESC
and how they depend on the Observing Strategy.


%---------------------------------------------------------------------

\input{Cosmology/lss.tex}

% --------------------------------------------------------------------

\input{Cosmology/wl.tex}

% --------------------------------------------------------------------

\input{Cosmology/photoz.tex}

% --------------------------------------------------------------------

\input{Cosmology/supernovacosmology.tex}

% --------------------------------------------------------------------

\input{Cosmology/lenstimedelays.tex}

% --------------------------------------------------------------------
