% ====================================================================

% LSST Observing Strategy White Paper

% Copyright 2015 The LSST Science Collaborations

% ====================================================================

\documentclass[11pt,headsepline,cleardoubleempty,twoside,openright]{scrbook}
% 11pt font
% Draw a line under the header
% Don't draw a line or print page numbers when a page is empty
% Pages are two-sided - so margins alternate
% Chapters start on the righthand side of the page

\usepackage{LSST_Observing_Strategy_White_Paper}

% ====================================================================

\begin{document}

\begin{titlepage}
\begin{center}

\vspace*{\stretch{3}}

{\Huge\bfseries\scshape
 Science Driven Optimization \\
 of the LSST Observing Strategy}

\vspace*{\stretch{2}}

% Version number, to be taken out for published draft

% \include{VersionDate}

% \begin{center}
% {\Large\bf Version 1.0\\
%
% \vspace*{\stretch{0.2}}
%
% November 2009}
% \end{center}

\vspace*{\stretch{2.5}}

{\Large  Prepared by the LSST Science Collaborations,}\\
\vspace*{\stretch{0.15}}
{\Large with contributions from the LSST Project. }\\
\vspace*{\stretch{1}}

\end{center}
\end{titlepage}

% --------------------------------------------------------------------

\tableofcontents

% --------------------------------------------------------------------
\clearemptydoublepage

\chapter*{Preface}
\addcontentsline{toc}{section}{Preface}
\markboth{}{}

\noindent This is a community white paper outlining various science
cases and the impacts that observing strategy will have on them,
quantified using the Metric Analysis Framework. We will describe
various strategies and tradeoffs that impact the observing cadence
(visit sequence), the current cadence baseline, and future directions
for the optimization of the  survey strategy. We aim to publish this
white paper on arXiv, and invite community feedback.

The timescale for producing this white paper, started before and
finished after the Observing Strategy workshop at the  August 2015
LSST Project and Community workshop, is many months.

The main points we will aim to convey in this white paper are as follows:

\begin{itemize}

    \item We have a pretty good idea of how we would deploy LSST:
    there is a baseline strategy and example cadences, with which it
    can be demonstrated that the data required for the promised
    science can be delivered.

    \item The baseline strategy can and will be optimized -- even small
    improvements can be significant. Most importantly, the strategy is
    not set in stone and it will evolve.

    \item The cadence optimization process will be as open and
    inclusive as technically possible. All stakeholders will
    participate in this process.

\end{itemize}

\raggedright{Project start: July 2015.}


% --------------------------------------------------------------------

\include{introduction}

% --------------------------------------------------------------------

\include{solarsystem}

% --------------------------------------------------------------------

\include{galaxy}

% --------------------------------------------------------------------

\include{astrometry}

% --------------------------------------------------------------------

\include{variablesandtransients}

% --------------------------------------------------------------------

\include{cosmology}

% --------------------------------------------------------------------

\include{supernovacosmology}

% --------------------------------------------------------------------

\include{deepdrilling}

% --------------------------------------------------------------------

\include{specialsurveys}

% --------------------------------------------------------------------

\include{tradeoffs}

% --------------------------------------------------------------------

\bibliographystyle{apj}
\bibliography{references}

% --------------------------------------------------------------------

\end{document}

% ====================================================================
