% ====================================================================
%+
% NAME:
%    section-name.tex
%
% ELEVATOR PITCH:
%    Explain in a few sentences what the relevant discovery or
%    measurement is going to be discussed, and what will be important
%    about it. This is for the browsing reader to get a quick feel
%    for what this section is about.
%
% COMMENTS:
%
%
% BUGS:
%
%
% AUTHORS:
%    David Nidever (@dnidever)
%    Knut Olsen (@knutago)
%-
% ====================================================================

\section{ The Magellanic Clouds }
\def\secname{MCs}\label{sec:\secname}

\noindent{\it David L. Nidever, Knut Olsen} % (Writing team)

% This individual section will need to describe the particular
% discoveries and measurements that are being targeted in this section's
% science case. It will be helpful to think of a ``science case" as a
% ``science project" that the authors {\it actually plan to do}. Then,
% the sections can follow the tried and tested format of an observing
% proposal: a brief description of the investigation, with references,
% followed by a technical feasibility piece. This latter part will need
% to be quantified using the MAF framework, via a set of metrics that
% need to be computed for any given observing strategy to quantify its
% impact on the described science case. Ideally, these metrics would be
% combined in a well-motivated figure of merit. The section can conclude
% with a discussion of any risks that have been identified, and how
% these could be mitigated.

%A short preamble goes here. What's the context for this science
%project? Where does it fit in the big picture?
The Magellanic Clouds have always had outsized importance for astrophysics.  They are critical steps in the cosmological distance ladder, they are a binary galaxy system with a unique interaction history, and they are laboratories for studying all manner of astrophysical phenomena.  They are often used as jumping-off points for investigations of much larger scope and scale; examples are the searches for extragalactic supernova prompted by the explosion of SN1987A and the dark matter searches through the technique of gravitational microlensing.  More than 17,000 papers in the NASA ADS include the words ``Magellanic Clouds'' in their abstracts or as part of their keywords, highlighting their importance for a wide variety of astronomical studies.

An LSST survey that did not include coverage of the Magellanic Clouds and their periphery would be tragically incomplete.  LSST has a unique role to play in surveys of the Clouds.  First, its large $A\Omega$ will allow us to probe the thousands of square degrees that comprise the extended periphery of the Magellanic Clouds with unprecedented completeness and depth, allowing us to detect and map their extended disks, stellar halos, and debris from interactions that we already have strong evidence must exist (REFS).  Second, the ability of LSST to map the entire main bodies in only a few pointings will allow us to identify and classify their extensive variable source populations with unprecedented time and areal coverage, discovering, for example, extragalactic planets, rare variables and transients, and light echoes from explosive events that occurred thousands of years ago (REFS).  Finally, the large number of observing opportunities that the LSST 10-year survey will provide will enable us to produce a static imaging mosaic of the main bodies of the Clouds with extraordinary image quality, an invaluable legacy product of LSST.

We propose two distinct mini-surveys to meet the goals of LSST Magellanic Clouds science:
\begin{itemize}
\item A mini-survey covering the 2700$\deg^2$ with $\delta < -60$ to the standard LSST single-exposure depth and to stacked depths of XXX, with cadence sufficient to detect and measure light curves of RR Lyrae stars 
\item A mini-survey covering $\sim$250$\deg^2$ of the main bodies of the Clouds with cadence sufficient to detect exoplanet transits and other variable objects; a subset of these images should be taken with seeing of $0.5\arcsec$, with stacked depth reaching the confusion limits in the Clouds
\end{itemize}

These surveys will support several important scientific goals:

Figure X shows a rough map of the proposed mini-surveys.  
% Need the figure and caption
These surveys will support several important scientific goals:
\begin{enumerate}
\item What are the stellar and dark matter mass profiles of the Magellanic Clouds?  Map extended disk, halo, debris, and streams.  Use streams as probes of total mass profile.  RR Lyrae give potential for three-dimensional stellar profile.
\item What is the satellite population of the Magellanic Clouds?  Discovery of dwarfs by DES and other surveys illuminating for understanding distribution of dark matter subhalos and how galaxies form in them (REFS)
\item What are the internal dynamics of the Magellanic Clouds?  Proper motions from HST and from the ground (REFS) have measured the bulk motions of the Clouds and have, in combination with spectroscopy, begun to unravel the three dimensional internal dynamics of the Clouds...
\item How do exoplanet statistics in the Magellanic Clouds compare to those in the Milky Way?  Lund calculation shows can measure transits of Jupiter-like planets, Clouds are lower metallicity environment
\item Identify and characterize the variable star and transient population of the Clouds.  Population studies, linking to star formation and chemical enrichment histories, etc, from Szkody et al. DD white paper.
\item Light echoes from supernovae and explosive events.  Echoes can give view of such events unavailable by any other means, ref. papers by Rest et al.
\end{enumerate}

% Convert list to short sections to provide background on these topics?

%David's text
Two main overarching science themes:
\begin{enumerate}
\item {\bf Galaxy formation evolution}: The study of the formation and evolution of the Large and Small
  Magellanic Clouds (LMC and SMC, respectively), especially their interaction with each other and the Milky Way.
  The Magellanic Clouds (MCs) are a unique local laboratory for studying the formation and evolution of
  dwarf galaxies in exquisite detail.  LSST's large FOV will be able to map out the three-dimensional
  structure, metallicity and kinematics in great detail.
\item {\bf Stellar astrophysics \& Exoplanets}:  The MCs have been used for decades to study stellar
  astrophysics, microlensing and other processes.  The fact that the objects are effectively all at a single
  known distance makes it much easier to study them than in, for example, the Milky Way.  LSST will extend
  these studies to fainter magnitudes, higher cadence, and larger area.
\end{enumerate}

Many different types of objects and measurements with their own cadence ``requirements'' will fall into
these two broad categories (with some overlap).  These will be outlined in the next section.

A very important aspect of the ``galaxy evolution'' science theme is not just the cadence but also the
sky coverage of the Magellanic Clouds ``mini-survey''.  A common misunderstanding is that the MCs only
cover a few degrees on the sky.  That is, however, just the central regions of the MCs akin to the thinking
of the Milky Way as the just the bulge.  The full galaxies are actually much larger with LMC stars
detected at $\sim$21$^{\circ}$ ($\sim$18 kpc) and SMC stars at $\sim$10$^{\circ}$ ($\sim$11 kpc) from their
respective centers.  The extended stellar debris from their interaction likely extends to even larger
distances.  Therefore, to get a complete picture of the complex strucure of the MCs will require a
mini-survey that covers $\sim$2000 deg$^2$.  At this point, it not entirely clear how to include this
into the metrics.  Note, that for the second science case this is not as much of an issue since the large
majority of the relevant objects will be located in the high-density, central regions of the MCs.


% static science vs. variables
% cadence vs. areal coverage
%  -want to cover whole SCP area to some necessary depth
%  -cadence for time-variable objects in full area or portion
%    what's the minimal # of visits (with decent phase coverage) to do this?  ask Kathy
%    Paula Szkody white paper on doing the full MCs for variables, linked to old MC cadence page
%   maybe a separate metric for spatial structure
%   -variabes:  ability to recover period, classify, 


% --------------------------------------------------------------------

\subsection{Target measurements and discoveries}
\label{sec:keyword:targets}

%Describe the discoveries and measurements you want to make.
%
%Now, describe their response to the observing strategy. Qualitatively,
%how will the science project be affected by the observing schedule and
%conditions? In broad terms, how would we expect the observing strategy
%to be optimized for this science?

\begin{enumerate}

\item Deep Color Magnitude Diagrams
%  -Deep CMDs, just a matter of number of visits
%  -do the full SMASH (and relevant DES area) with full spatial coverage, at least to SMASH depths, smaller
%  number of epochs, ~5 sigma at gri~25
% Knut thoughts: I think we want to make sure that we get 1 mag below old turnoff out to 100 kpc in ugriz with 10sigma precision, i.e. ugriz~25


\item Proper Motions
%-Proper Motions, cadence not as much of an issue, just more epochs
%  bulk proper motion
%  LMC spiral motion, streaming motions
%  internal velocity dispersion

In the last decade work with $HST$ has been able to measure the bulk tangential (in the plane of the sky) velocities ($\sim$300 km/s) of the
Magellanic Clouds (Kallivayalil et al.\ 2016a,b,2013) and even the rotation of the LMC disk (van der Marel \& Kallivayalil 2014).
Gaia will measure precise proper motions of stars to $\sim$20th magnitude which will include the Magellanic red giant branch stars.
LSST will be complementary to Gaia and measure proper motions of stars in the $\sim$20--24 mag range that includes Magellanic main-squence stars
which are far more numerous than giants, and, therefore, more useful for mapping extended stellar structures.
The LSST 10-year survey proper motion precision will be $\sim$0.3--0.4 mas/yr at LMC main-sequence turnoff at r$\approx$22.5--23.  This
will allow for accurate measmurent of proper motions of {\em individual stars} at the $\sim$5$\sigma$ level.

Besides measuring kinematics, the LSST proper motions can be used to produce clean samples of Magellanic stars 
clean samples of background galaxies (no proper motions) and 
this is commonly done with $HST$ data of 

In addition, LSST proper motions can be used to improve star/galaxy separation which is quite significant for faint, blue Magellanic
main-sequence stars.

streaming motions
%can we do individual LMC stars with LSST, or small groups?

% SRD says want 0.2 mas/yr accuracy over the course of the survey
% 0.2 mas/yr at r=20.5 (similar to Gaia)
% ~0.25 mas/yr at r=22
% ~0.3 mas/yr at r=22.5 LMC turnoff
% ~0.4 mas/yr at r=23
% 1 mas/yr for r=24
% See Figure 21 from Ivezic et al. (2012) or slide 46 of overview-sci-reqs.pdf
%have the astrometric precision to measure the proper motions of individual
%Proper motion cleaning to find the giants?? gaia does that already
%lsst can use proper motion cleaning to do star/galaxy separation as well
%
%The 
%The Magellanic Clouds have a large tangential velocity (in the plane of the sky) that has been 
%Gaia will be able to see the bright stuff, need lsst to get the MSTO
%gaia/lsst synergy
%
%metric that calculates the proper motion accuracy of LMC MSTO stars at r=23 and calculates the sigma-level of
%the proper motion measurement (2.0 mas/yr / sigma_pm ).



%\item Parallaxes
%-Parallaxes, also mostly a function of nubmer of epochs
%  bulk distances
%  internal distance spread
%  -probably not get parallaxes at MC distances, but could do foreground rejection
  
\item Variable stars

%-Variables, RR Lyrae, Cepheids might be too bright, dwarf cepheids/scuti good, many more of them.
%   especially good for getting the 3D structure (out to large distances) of the MCs
%   -eclipsing binaries (get very accurate distances, see OGLE paper), pulsating WDs, CVs, T Tauri stars
% novae, supernovae  in Paula's white paper

% use MCs are a way to get templates for variable sources that LSST will detect all over the sky
% look at variable group metrics
  
  
\item Transients
%-Transients, dwarf novae
% Mike Lund will work on some text for this.  Also did work on cadence considerations
% for detecting periodic objects in general (periodogram purity function).
  
\item Transiting Exoplanets
% -Transiting planets, Mike Lund
% -need deep drilling to have any hope of finding them
% -best between ~0.8-1.6 Msun to detect exoplanets, not too faint
% -can't get ingress/egress but can detect the dip and periodicity
%   that's enough to characterize
% -challenging to do follow-up because they are quite faint and too many (?)
% -cadence?  cover all timescales properly
% -what's the expected period distribution
See Lund et al.\ (2015) and the exoplanet discussion in previous part of this paper.
% PASP
  
%\item Astrometric binaries
%-Astrometric binaries

\item Light-echoes
% it's a surface-brightness issue, can see fainter things with LSST than MOSAIC/DECAM
% could trace out lightcurve if more epochs  
  
\item Gyrochronology
%-Gychronology, need to get periods of the dwarfs, gives age information
%  -gyro periods are ~10 days at 1 Gyr and ~30 days at 10 Gyr
  
\item Interstellar scintillation
% run in movie mode for 1-2 nights to find missing molecular gas (H2)
% https://github.com/LSSTScienceCollaborations/ObservingStrategy/issues/68

% Legacy survey
% use the best seeing to get great data on the MC main bodies
  
%\item Astroseismology
%-Astroseismology, dwarfs/giants, giants vary by a couple percent and on "longer" timescales, but
%    probably too bright for LSST, OGLE probably has best data for those. however LSST might be able to do
%    asteroseismology of giants to larger distances, measure masses/ages of halo giants!
%    dwarfs are harder because they vary less and need more higher frequency observations

\end{enumerate}

% --------------------------------------------------------------------

\subsection{Metrics}
\label{sec:keyword:metrics}

Quantifying the response via MAF metrics: definition of the metrics,
and any derived overall figure of merit.

% metric on surface brightness limit in different parts of the sky
% to MC structure
% could use metric for how much of the Besla stellar debris we can detect
%  even just that region of the sky covered

\item  Magellanic Clouds proper motion metric.
It would be useful to calculate a metric that's the significance level ($\sigma$-level) of the proper motion measurement of one Magellanic MSTO star (r=23 mag).
This would be $\sim$2.0 mas/yr / $\sigma_{\rm pm}$.
%  the precision proper motion metric that 
%  Metric that calculates the proper motion accuracy of LMC MSTO stars at r=23 and calculates the sigma-level of
%the proper motion measurement (2.0 mas/yr / sigma_pm ).

\item Another useful metric would be to calculate the surface brightness limit of the Magellanic structures using MSTO stars.

\item Metric of how much of the expected Magellanic debris/structure (from Besla et al.\ 2012 model) we can detect.  This would depend mostly
on the area covered, but we could also use the surface brightness limit (calculated above) directly.


% --------------------------------------------------------------------

\subsection{OpSim Analysis}
\label{sec:keyword:analysis}

OpSim analysis: how good would the default observing strategy be, at
the time of writing for this science project?


% --------------------------------------------------------------------

\subsection{Discussion}
\label{sec:keyword:discussion}

Discussion: what risks have been identified? What suggestions could be
made to improve this science project's figure of merit, and mitigate
the identified risks?


% ====================================================================

\navigationbar
