\chapter[Special Surveys]{Special Surveys}
\def\chpname{specialsurveys}\label{chp:\chpname}

Chapter editors:
\credit{dnidever},
\credit{knutago}.

% Confirmed leads for LMC/SMC: Knut Olsen, David Nidever

% Confirmed leads for special surveys:

% \section*{Summary}
% \addcontentsline{toc}{section}{~~~~~~~~~Summary}
%
% Executive summary goes here, highlighting the primary conclusions from
% the chapter's science cases. This should be abstract length, no more:
% say, 200 words.

% --------------------------------------------------------------------

\section{Introduction}
\label{sec:specials:intro}

The four main LSST science themes, as defined by the Science Book,
drive the design of LSST's main Wide-Fast-Deep survey.  However, it
has always been recognized that many important scientific projects,
including some that are highly relevant to LSST's main science themes,
are not well served by the areal coverage and/or cadence constraints
placed on the WFD survey.  To this end, the LSST Project set aside a
nominal 10\% of the observing time to serve what are collectively
called ``special surveys.''

Projects that
will certainly make use of this $\approx10\%$ time (that is not dedicated to the WFD
survey) include the Deep Drilling fields and the Galactic Plane surveys,
as well as any survey wishing to
observe at declinations below $-60^\circ$, such as the Magellanic
Clouds.  These special programs have the potential to
heavily oversubscribe the nominal 10\%
of time assigned to them.  It is of thus critical importance for these
programs to define compelling science cases, clearly justify their
observing requirements, and derive metrics to quantify the performance
of a given schedule for the program. This chapter provides a venue for
such investigations.

A minimal set of 4 ``extragalactic'' Deep Drilling Fields have been
included in many of the \OpSim runs to date, including the baseline cadence simulation, \opsimdbref{db:baseCadence} (\autoref{chp:cadexp}), and have
been evaluated in various science sections throughout this paper.
As described in \autoref{sec:intro:timeline}, there will be a further call for Deep Drilling Field proposals, and so we defer discussion of such observing programs to that activity. In this chapter we collect together various ideas for additional special surveys or otherwise unusual observing programs, and the discussion of their metric evaluation.

In \autoref{sec:solar_system_specials} a number of different concepts for special surveys that would support solar system science is described. This is followed by an extensive discussion of short exposure observing at twighlight (\autoref{sec:shortexp}). If short exposure times will indeed be a possibility, it won't just be solar system science that benefits: an example of a small program that could make good use of such a capability is the open star cluster special survey proposed in \autoref{sec:M67_special}. We expect the ideas in this chapter to be developed (or discarded) over the time before the survey starts: this white paper aims to provide a (potentially temporary) home for them during that period.


% Add sections below, one science investigation per section, one
% section per file.

% --------------------------------------------------------------------

% PJM: commented out for now, for lack of content:
% \input{SpecialSurveys/deepdrilling.tex}

% --------------------------------------------------------------------

% PJM: This is currently in the Magellanic Clouds Chapter, but
% could be moved back here soon...
% \input{SpecialSurveys/magcloudsurvey.tex}

% --------------------------------------------------------------------

\input{SpecialSurveys/solar_system_specials.tex}

% --------------------------------------------------------------------

\input{SpecialSurveys/short_exposures.tex}

% --------------------------------------------------------------------

\input{SpecialSurveys/M67_special.tex}

% --------------------------------------------------------------------
