% ====================================================================
%+
% SECTION:
%    tde.tex  
%
% CHAPTER:
%    transients.tex  
%
% ELEVATOR PITCH:
%    Tidal disruption events (TDEs) are the disruptions of stars by supermassive black holes.
%    They can produce flares in the optical and UV (sometimes accompanied by X-ray and radio emission as well).
%    These flares can be used to reveal the properties of otherwise quiescent SMBHs and to study accretion physics.
%    TDEs are rare, LSST will allow the first statistical sample of such events.
%
% COMMENTS:
%
%
% BUGS:
%
%
% AUTHORS:
%   Iair Arcavi (@arcavi)
%
% ====================================================================

\section{Tidal Disruption Events}
\def\secname{tdes}\label{sec:\secname} % For example, replace "keyword" with "lenstimedelays"

\credit{arcavi} % (Writing team)

A star passing close to a supermassive black hole (SMBH; $M\gtrsim10^{6}M_{\odot}$)
will be torn apart by tidal forces. For certain ($\lesssim10^{8}M_{\odot}$)
black hole masses, the disruption will occur outside the event horizon
and will be accompanied by an observable flare (Hills 1975, Rees 1988).
Such flares can be used to study inactive SMBHs, which are otherwise
inaccessible beyond the nearby ($\lesssim100$ Mpc) universe. 

We are now building our understanding of how the observational
properties of TDEs are affected by the SMBH. Theory claims to provide
such a connection (e.g. Lodato et al. 2009, Guillochon et al. 2014),
but uncertainties in the physics of the disruption, subsequent accretion
and emission mechanisms are currently topics of debate (e.g. Strubbe \&
Murray 2015), and new models are vigorously being developed (e.g.
Piran et al. 2015; Hayasaki et al. 2015; Svirski et al. 2015; Bonnerot
et al. 2016). 

TDEs are rare ($\sim10^{-5}-10^{-4}$ events per galaxy per year; Wang \& Merritt 2004; 
Metzger \& Stone 2016), and until recently, TDE candidates were discovered 
mostly in archival data (e.g. Donley et al. 2002, Gezari
et al. 2006, Esquej et al. 2007). However, wide-field transient
surveys have recently started discovering TDEs in real time.

Generally, two types of TDE candidates have been well observed:
\begin{enumerate}
\item High energy TDEs - The prototype of which is Swift J1644 (refs), and two
other events are known. These events display emission in gamma- and X-rays as well as
in the radio, but are not detected in the optical. 
\item Optical-UV TDEs - The prototype of which is PS1-10jh (Gezari et al. 2012).
About $8$ other events are known (Chornock et al. 2014, Arcavi et al. 2014, Holoien et al. 2016a,b).
Some events were detected also in the X-rays and radio (in addition to the optical and UV),
but the X-ray and radio signatures are different than those of the high energy TDE candidates.
\end{enumerate}

It is still not clear whether both of these classes of transients are TDEs, 
and if so, why they are so different form eachother. One option raised is 
that some TDEs may launch jets, which when directed towards us, appear as
the high energy events, but otherwise appear as the optical-UV events.
It is still not clear if this is the case (e.g. van Velzen et al. 2011).

Here we focus on the second type of TDE candidates, which is the relevant
class for LSST, since they can be discovered in the optical. However,
multi-wavelength coordinated observations are required in order to better 
understand the connection between the two types of candidates.

The first well-sampled TDE of the optical+UV class was PS1-10jh (discovered by Pan-STARRS;
Gezari et al. 2012). Arcavi et al. (2014) later presented three new TDE candidates
from PTF and one discovered by ASAS-SN, all with similar properties
as PS1-10jh. These events exhibit blue colors, broad light curves,
peak absolute magnitudes of $\sim-20$ and a $\sim t^{-5/3}$ decay
at late times. This decay law has been suggested as a unique signature
of accretion-powered TDE light curves (Rees 1988, Evans \& Kochanek
1989, Phinney 1989). Early-time deviations from the $t^{-5/3}$ rate
can be used to constrain the density profile of the disrupted star
(Lodato et al. 2009, Gezari et al. 2012). Late-time deviations would
test the accretion power-source hypothesis altogether. 

The spectral signatures of these TDEs are still a puzzle. PS1-10jh displayed
only He II emission lines, lacking any signs of H. Some of the Arcavi
et al. (2014) sample, however, do display H emission (Fig. \ref{fig:tdes}).
In fact, a continuum of H / He emission ratios for this class of transients
is being revealed, and is now a focus of theoretical modelling (Strubbe
\& Murray 2015; Roth et al. 2015).

The second recent discovery relating to this new sample concerns their
host galaxies (Arcavi et al. 2014; French et al. 2016), most of which are
post-starburst galaxies. These galaxies show little or no signs of on-going star formation, but their significant A stellar populations indicate that star formation ceased abruptly a few hundred Myr to a Gyr ago (Dressler \& Gunn 1983). Galaxies with these characteristics often show signs of recent galaxy-galaxy mergers (Zabludoff et al. 1996), which produced the starburst and evolved the bulge. Optical-UV TDEs are intrinsically over abundant in post-starburst galaxies by a factor of $\sim30-200$ (depending on the characteristics of the galaxy; French et al. 2016). The reason for the strong preference of TDEs for post-starburst galaxies has still not been determined.

LSST's contribution to TDE studies will be substantial. van Velzen et al. (2011) estimate that LSST could discover approximately 4000 TDEs per year. The main drivers for studying TDEs with LSST are:
\begin{itemize}
\item Measuring black hole masses: This involves fitting models to TDE light curves. It is also relevant to correlate these measurements with host galaxy properties (mass, bulge/disk decomposition). 
\item Constraining galactic dynamics by measuring the TDE rates as functions of black hole mass and galaxy types. 
\item Characterizing TDE emission signatures. 
\end{itemize}

A metric is required for measuring how well TDEs can be identified and distinguished from supernovae and active galactic nuclei. In general, we expect TDEs to:
\begin{itemize}
\item Be located in the center of their host.
\item Display approximately constant blue (few $10^4$K) colors.
\item Evolve slowly (weeks-months).
\item Not show past AGN-like variability.
\item Preferentially peak around mag -20.
\item Preferentially be hosted in a post-starburst galaxy.
\end{itemize}
These criteria are based on our current knowledge of optical TDEs, which is still in its early stages. The field is rapidly evolving, and it is possible that new observations will change the current picture of TDE emission. 
This metric is probably best combined with those discussed in Section \ref{sec:SNtransients} for identifying supernovae.

A second metric is required to asses the accuracy with which the black hole mass can be constrained from the TDE light curves. This metric can be based on existing theoretical models to fit simulated TDE light curves (such as TDEFit; ref. Guillochon). 


