% --------------------------------------------------------------------

\chapter[Eruptive and Explosive Transients]{Eruptive and Explosive Transients}
\def\chpname{transients}\label{chp:\chpname}

% \noindent {\it
% Mike Lund, Ashish Mahabal, Stephen Ridgway, Lucianne Walkowicz, Rahul Biswas, Michelle Lochner, Jeonghee Rho, Eric Bellm, Federica B. Bianco...
% }

Chapter editors:
\credit{ebellm},
\credit{fedhere},

Contributing Authors:
\credit{AshishMahabal},
\credit{StephenRidgway},
\credit{ohadshemmer},
\credit{paulaszkody},
\credit{nathansmith},
\credit{tommatheson}

% --------------------------------------------------------------------

\section{Introduction}


Explosive and eruptive transients include a diverse ensemble of
object, diverse both physically and phenomenolocigally. Transients
such as novae, supernovae (SNe), and gamma-ray bursts (GRBs) probe the
final stages of stellar evolution, Tidal Disruption Events (TDEs), GRBs, Cataclysmic Variables (CVs) give us the
opportunity to study compact and binary objects, massive star
eruptions allow us to understand mass loss mechanisms, and chemical
enrichment. Transients impact the evolutionary history of the
Universe. The brightest transients (GRBs, TDEs, SNe) are light beams
that can be seen over cosmic distances, and some transients (Type Ia
SNe in the first place) are cosmological tracers.  Cadence choices
will determine LSST's ability to discover, classify, and characterize
these events. However, different types and time scales of phenomena
will benefit from different sampling strategies---sometimes
significantly different, and at times orthogonal.  Competing
objectives described in this chapter are at the heart of LSST
observing strategy and cadence design.

When evaluating a particular observation or series of observations in
light of how they perform for a specific science case, it may be
helpful to think of metrics as lying along a continuum between
discovery and characterization. Discovery requires a minimum amount of
information to recognize an event or object as a candidate of
interest, which necessarily involves some level of bare-bones
characterization; rich characterization, on the other hand, implies
that an event may not only be recognized as a candidate of interest,
but basic properties of the event or object may be determined from the
LSST observations (including but not limited to the classification of
the event). The level of characterization accessible through LSST data
will of course evolve with time for all the transients that have
characteristic time scales longer than a few days.

Here we foucs on cadences feasible within the main survey.
Many science objectives related to transients may be better, and possibly only,
achieved within the DDF program.
Transient events may benefit from substantial temporal sampling
(matched to the time constant of the event) with color information
(perhaps contemporaneous) to support characterization and
classification, obtained over the limited duration of interest.
Transient events slower than $\sim$ weeks may be adequately sampled by
a uniform LSST cadence.  Faster events may require special scheduling
strategies.  For some event types, LSST can only be expected to
provide a discovery service, and followup will necessarily be
performed elsewhere.  For some events, for example detecting LIGO
counterparts, a serendipitous discovery is extremely unlikely, but
enabling a TOO program would provide the opportunity for LSST to
contribution significantly to this science.

%The interpretation of a given metric along this continuum has
%implications for the subsequent action and analysis required,
%particularly as regards possible follow-up observations with other
%facilities.

We consider a non-exhaustive set of ``astronomical transients'' in the
paragraphs that follow.

\subsection{Targets and Measurements}
\label{sec:\chpname:targets}

%The class of transients includes a heterogeneous assortment of objects
%and phenomena.
The table below is a \emph{non-exhaustive} list of
phenomena to which we are referring as \emph{Eruptive and Explosive
  transients} in this document.

\begin{center}
  \begin{tabular}{| p{2.5cm} | p{6cm} | l | l | l |}
    \hline

    Transient Type & Science drivers & Amplitude & Time Scale & Event Rate\\
\hline

Flare stars & Flare frequency, energy, stellar age & large & min & ?\\

X-ray Novae & Interacting binaries, stellar evolution, SN progenitors, nuclear physics & large & weeks & ?\\

Cataclysmic variables & Interacting binaries, stellar evolution, compact objects & large & min - days & ?\\

LBV variability & Late stages stellar evolution, Mass loss, SN progenitors & large & weeks-years & ?\\

Massive star eruptions & Late stages stellar evolution, Mass loss, SN progenitors & extreme & weeks-years & rare\\

SN & stellar evolution, feedback, chemical enrichment, cosmology & extreme & days - months & very common\\

GRBs & SN connection, stellar evolution, cosmology & extreme & min - days & ?\\

TDEs & Massive BH demographic & large & months & ?\\

LIGO detections & EM characterization & unknown & unknown & ?\\

Unknown & Discovery & unknown & unknown & rare\\

 \hline \end{tabular}
\end{center}

WE SOULD ADD A FIGURE OF THE PHASE SPACE + EVENT RATE


{\bf What else?}


%templates and stacks

%confusion 

%


% --------------------------------------------------------------------


\subsection{Transient time scales}

For very short lived phenomena (stellar flares, CV outbursts, GRBs,
LIGO events) the main function of LSST will be to provide discoveries
and/or simple characterization.  Followup to discovery/identification,
if required, will surely take place elsewhere.

LBV's and massive star eruptions (such as the 
Gap transients??

SN fall in an intermediate time range.  LSST will provide
multiple visits in multiple filters during the typical SN duration
(months).  This sampling may be insufficient for many science
objectives.  However, moderate, and feasible, changes to LSST
observing strategy, may enhance the sampling for part of the sky part
of the time, greatly improving the usefulness of SN observations. 
Metrics that assess the discovery rate of SN are included in the
cosmology chapter. Here we are interested in assessing the ability of
LSST to discriminate SN from other transients, SN subtypes from one
another, and to identify particularly interesting SNe: for example
those that show signature of shock break-out, or companion interaction
in the early light curve, and would be candidates for \emph{flash
  spectroscopy} follow-up (ref here). In addition metrics that
quantify LSST's ability to constraint SN physics in a statistically
large sample of SN are needed (offered hopefully).


**** STILL WANT THIS? CAN WE DO IT?
TDEs, whose fading time-scale is much more gradual
(over weeks to months) than the rise time-scale and we want to asses -
through a metric - how many are detected while still on the
rise. Ref. Science Book: 10.6.1. Also ref. recent papers.

A phenomenon that falls under this document's definition of explosive
transients, but is not discussed here, is AGN. AGN are likely to be
sampled with sufficient resolution by a uniform LSST cadence and they
are discussed in Chapter~\ref{chp:agn}.

In addition we hope that LSST will provide a wealth of serendipitous
discoveries of yet-to-be-observed transients.  An ideal transient
discovery survey would include heavy coverage of all time scales. LSST
will cover longer time periods well, but will have to make some
choices of emphasis in coverage of shorter time-scales. (metric
here??)


In the paragraphs that follow we will use case studies to asses:

\begin{itemize}
\item
  The discerning power of a given LSST observing cadence.
  While early classification of transients is extremely important for
  transient science with LSST, it is also an incerdibly difficult task,
  on which many astronomers and statisticians are working. It is unfeasible,
  and beyond the scope of this work to attempt a naive classification metric.
  However a crucial input to follo-up strategy design is the early
  identification of young, fast evolving transients. We identify a region of
  the phase-space of rising speed and color that is characteristif of a
  variety of transients in their early phases and assess the ability of LSST's
  cadences to place transients within this phase-space. (Eric, Stefano, Iair, Fed...)
  %For this topic we will address the ability to distinguish LBVs, SNe, and TDEs.
\item
  The ability to identify an object of interest for follow-up and
  trigger prompt follow-up observations. For this topic GRBs are used as
  a case study. ***ERIC
\item
  The insight that a cadence gives into a single transient class. The
  case study will be
  **massive star eruptions if Nathan gets it done
  **CVs if Laura gets it done
\item 
  The statistical constraints to a transient class that can be obtained
  over the course of the LSST survey, from the LSST survey data alone
  (assuming a successful classification). SN Ia early interaction
  signatures or IIb shock break-out (Fed).
\item
  The ability of LSST, given a cadence, to provide insightful coverage
  for long duration transients. This metric favors cadences with small
  seasonal gaps, which allow us to gather complete color evolution of
  slow evolving transients, such as LBV and TDEs. (Ryan)

\end{itemize}

In this chapter we focus on LSST's potential to advance the science of
eruptive and explosive transients; the use of SNe for cosmology is
discussed in \autoref{chp:cosmo}.


% --------------------------------------------------------------------

\subsection{Metrics}
\label{sec:\chpname:metrics}

\begin{center}
\begin{tabular}{| p{5cm} |p{10cm} |}
\hline Metric & Description\\ \hline SNe & Number of events adequately
sampled\\ Serendipity & Histogram of median visit series length vs
maximum visit spacing within the series\\ \hline \end{tabular}
 \end{center}


%%WHO WROTE THIS? I AM NOT SURE WHAT THIS METRIC DESCRIPTION MEANS
The metrics for SNe will be highly specialized and based on the best
available understanding of SN light curve analysis and the expected
event population.

%%WHAT DOES THE PARAGRAPH BELOW MEAN?
The suggested metric set for serendipity is based on the simple-minded
idea that a novel transient will be characterized by a band-limited,
finite waveform, and that a useful observation series will consist of
a series of samples extending over the duration of the event, with at
least critical sampling of the fastest variations.  Since for some
event durations the number of useful time series will be small, it may
be useful to look not at the median length, but the median length of a
subset size preselected as possibly useful (e.g. the$10^3$ longest
series).

\citet{2015arXiv150803175L} discuss
three metrics that have been incorporated into the MAF. Two of these
metrics deal explicitly with time variable behavior: a) observational
triplets, and b) detection of periodic variability. While in this
chapter w do not deal with periodic phenomena, the first of these
metrics is useful to generically assess the performance of LSST for
transients of all types. More specialized transients metrics can be
obtained by simply using a more realistic light curve shape. A metric
that uses input ASCII files is provided, to enable the assessment of
discovery and characterization of an arbitrarily shaped transient.

\subsubsection{Observational triplets (TripletMetric)}

This metric provides a means of evaluating whether a transient event
on some time scale of interest has been detected, by testing for a
sequence of three observations. The object must be detected in
quiescence, followed by two subsequent detections above some
threshold; this sequence of observations allows the magnitude of the
change to be measured, as well as its time scale.

This metric may be used for a variety of astrophysical phenomena, in
particular transient events on variable objects (e.g. novae, stellar
flares), in that it is general with respect to the amplitude of the
brightness variation as well as the time scale of said change. The
requirement of a detection prior to outburst does constrain it to
objects that have already been detected in quiescence (in other words,
not necessarily ``true'' transients), although there may be some cases
where this is not the case (e.g. a supernova occurring on a previously
detected galaxy). In practice, the time lapse between the first and
second and second and third observations must be comparable (between
10$^2$ and 10$^5$ seconds) for discovery. This metric may be
calculated for a given OpSim run and then further reduced to a
histogram in logarithmic time bins; the minimum number of bins to
construct an interesting sample of objects is source-dependent.

\subsubsection{Transient Metric (transientAsciiMetric)}

Calculate what fraction of transients would be detected using an ASCII
input file for the light curve.

\subsection{Proposed Metrics}

The following is a raw list of metric ideas; these need specificity
and further description.

The triplet metric may also be altered to include filter constraints,
such that the triplets are drawn from a single filter or subset of
filters.

Color evolution constraint: triplets of observations in a specific
color (really requirement of two triplets in multiple filters)

  2D Histogram of delta t?s between observations constituting a triplet

Histogram of median visit series length vs maximum visit spacing
within the series

Number of events adequately sampled

% --------------------------------------------------------------------

\subsection{OpSim Analysis}
\label{sec:\chpname:analysis}

Analysis shows that current simulations provide  poor coverage in any one filter for transient events longer than a deep drilling session ($\sim$30 minutes) and shorter than $\sim$ weeks.

Simulated performance for SN observations must be analyzed for both
main survey and mini-survey (deep drilling) productivity.  It is
considered that current simulated schedules give inadequate
performance for SN science.



% --------------------------------------------------------------------

\subsection{Discussion}
\label{sec:\chpname:discussion}

Community studies are providing improving SNe metrics, and continuing
communication between the SN and LSST communities is essential to
tuning the observing strategy to deliver the SN time series that are
needed and possible.

Improving LSST science return for SNe will also improve sampling of
all transients with similar or somewhat shorter characteristic times.
Non-uniform survey strategies (rolling cadence) can significantly
improve the LSST performance for faster transients.  Interpretation of
multiple filters for novel events may be powerful, or problematic,
since color may be uncertain.

Some insight into fast transients may be available from image pairs or
triples (as opposed to more complete series).  These include the pair
of images in a visit - which could be useful in studying the rise time
of an extremely fast event.  This includes the characteristic grouping
of visits (typically 0.5 to 1.0 hour separation) planned for purposes
of identifying asteroids.  It also includes fortuitous multiple
sampling due to field overlap, providing additional sampling, which
may be random or systematic, depending on the scheduling, on a time
scale of minutes to hours.  The sampling benefits of this fortuitous
overlap have not yet been investigated.


\navigationbar

% --------------------------------------------------------------------

\input{cv.tex}

% --------------------------------------------------------------------

\input{sn.tex}

% --------------------------------------------------------------------

\input{grb.tex}

% --------------------------------------------------------------------

\input{gw.tex}
