% --------------------------------------------------------------------

\chapter[Variable Objects]{Variable Objects}
\def\chpname{variables}\label{chp:\chpname}

Chapter editors:
\credit{AshishMahabal},
\credit{lmwalkowicz}.

% \noindent {\it
% Mike Lund, Ashish Mahabal, Stephen Ridgway,
% Lucianne Walkowicz, Rahul Biswas, Michelle Lochner,
% Jeonghee Rho, Eric Bellm...
% }

% --------------------------------------------------------------------


\section{Introduction}

Variable objects are defined as those that exhibit brightness changes, either periodic or non-periodic, which are detected in quiescence and non-destructive to the object itself. Variable objects span a wide range in timescale-of-interest (sometimes even within a single class of objects), and so different science cases benefit from different sampling strategies. These strategies may be significantly disparate from one another, sometimes even mutually exclusive; competing objectives described in this chapter and the next are therefore at the heart of LSST observing strategy and cadence design.

Below we develop a number of key science cases for LSST studies of variable objects, associating them with related metrics that can be used within the Metrics Analysis Framework (MAF) to understand the impact of a given survey strategy realization on the scientific results for that case. The science cases outlined are by no means exhaustive, but rather are motivated by providing key quantitative examples of LSST's performance given any particular deployment of survey strategy. The authors encourage community contribution of similar cases, where the scientific outcome can be quantified using specific metrics. 



%When evaluating a particular observation or series of observations in
%light of how they perform for a specific science case, it may be
%helpful to think of metrics as lying along a continuum between
%discovery and characterization. Discovery requires a minimum amount of
%information to recognize an event or object as a candidate of
%interest, which necessarily involves some level of bare-bones
%characterization (upon which said recognition is based); rich
%characterization, on the other hand, implies that an event may not
%only be recognized as a candidate of interest, but basic properties of
%the event or object may be determined from the observation (e.g.
%including but not limited to classification of the event). The
%interpretation of a given metric along this continuum has implications
%for the subsequent action and analysis required, particularly as
%regards possible follow-up observations with other facilities.

%Target types are here grouped in subsections by variability
%characteristics, but as will be seen, this does not mean that all
%targets in a group require a common cadence, since the times scales
%may vary dramatically.  Acquiring suitable data for a wide range of
%time scales presents a fundamental problem for LSST, since the
%available $~$800 visits to a field over the survey cannot be deployed
%so as to usefully sample all time scales at all times.  This fact
%leads to the concept of a non-uniform survey, in which parts of the
%sky are visited more frequently part of the time.  The merits of such
%options must be traded against the benefits of a more uniform survey
%strategy.

\begin{center}
\begin{tabular}{| l | p{8cm} |l | l |}
\hline Periodic Variable Type & Examples of target science & Amplitude & Timescale\\
\hline
RR Lyrae & Galactic structure, distance ladder, RR Lyrae properties&  large &  day \\
Cepheids & Distance ladder, cepheid properties&  large &  day \\
Long Period Variables & Distance ladder, LPV properties & large  &  weeks \\
Short period pulsators & Instability strip, white dwarf interior properties, evolution&  small & min  \\
Periodic binaries & Eclipses, physical properties of stars, distances, ages, evolution, apsidal precession, mass transfer induced period changes, Applegate effect &  small &  hr-day \\
Rotational Modulation & Gyrochronology, stellar activity& small  &  days \\
Young stellar populations & Star and planet formation, accretion physics & small  &  min-days \\
 \hline \end{tabular}
 \end{center}

\subsection{Description of Relevant Metrics}
\label{sec:keyword:variablemetrics}

Despite the range in scientific motivation for the cases presented below, there are some common metrics that are widely applicable (or may be combined in a variety of ways with other metrics to suit a variety of applications). 

%\subsection{Metrics}
%\label{sec:keyword:metrics}

\begin{center}
\begin{tabular}{| p{5cm} |p{10cm} |}
\hline Metric & Description\\
\hline
Eclipsing/transiting system discovery & Fraction of discoveries vs fractional duration of eclipse\\
%Transiting exoplanets (depth dependent) & Fraction of discoveries vs fractional duration of eclipse\\
Phase gap & Histogram vs period of the median and maximum phase gaps achieved in all fields\\
Period determination (period dependent) & Fraction of targets vs survey duration, for which the period can be determined to 5-sigma confidence\\
Period variability (period dependent) & Fraction of targets vs survey duration, for which a period change of 1\% can be determined with 5-sigma confidence\\
  \hline \end{tabular}
 \end{center}

The ability to identify that an object is periodic, and to correctly determine that object's period, are widely applicable measures of discovery. In the case of regular variables (as outlined below), these two measures together can uniquely identify a population. Other kinds of periodic systems (transiting planets for example) also require a measurement of periodicity, but have a much wider range of relevant periods, and looser requirements on the strictness of that periodicity. 

Lund et al. (2015; \url{http://arxiv.org/pdf/1508.03175.pdf}) discuss three metrics that have been incorporated into the MAF. Two of these metrics deal explicitly with time variable behavior: a) observational triplets, and b) detection of periodic variability. 

\subsubsection{Periodogram purity function (PeriodicMetric)}
This metric calculates the Fourier power spectral window function of each field (Roberts et al. 1987) as a means of quantifying the completeness of phase coverage for a given periodic variable. The periodogram purity is defined as 1 minus the Fourier power spectral window function; in the perfect case, all power is concentrated in a delta function at the correct frequency, and is zero elsewhere. As power ``leaks'' away from the correct frequency as a consequence of discrete, non-ideal data sampling, the periodogram becomes more structured. For the purposes of MAF metrics, which are designed to quantify performance as a single number, the periodogram purity is quantified as the minimum value away from the correct frequency. 

\subsubsection{Phase Gap Metric (PhaseGapMetric)}
Histogram of the median and maximum phase gaps achieved in all fields

\subsubsection{Period Deviation Metric (PeriodDeviationMetric)}

This metric computes the percent deviation of recovered periods from pure sine variability.

%\subsection{Proposed Metrics}

%The following is a raw list of metric ideas; these need specificity and further description. 

%FWHM of the window function (to quantify sampling)

%Maximum hour angle difference 

%Fraction of discoveries vs fractional duration of eclipse

%Fraction of targets vs survey duration, for which the period can be determined to 5-sigma confidence

%Fraction of targets vs survey duration, for which a period change of 1$\%$ can be determined with 5-sigma confidence


% --------------------------------------------------------------------

\input{periodicpulsators.tex}

% --------------------------------------------------------------------

\input{planets.tex}

% --------------------------------------------------------------------

\input{rotationalvariables.tex}

% --------------------------------------------------------------------

\input{youngstars.tex}

% --------------------------------------------------------------------
