% --------------------------------------------------------------------

\chapter[Variables and Transients]{Variable and Transient Sources in the Galaxy and Beyond}
\def\chpname{vartrans}\label{chp:\chpname}

\noindent {\it
Mike Lund, Ashish Mahabal, Stephen Ridgway, Lucianne Walkowicz, Rahul Biswas, Michelle Lochner,
Jeonghee Rho...
}

% --------------------------------------------------------------------


Confirmed leads for Stellar variability and Fast Stellar Transients:
Mike Lund, Ashish Mahabal, Stephen Ridgway, Lucianne Walkowicz

Confirmed leads for cosmological SNe: Rahul Biswas, Michelle Lochner,
Jeonghee Rho

Confirmed leads for extragalactic transients: Ashish Mahabal

Confirmed leads on rolling cadence: Stephen Ridgway

% --------------------------------------------------------------------

\section{Introduction}

The observation of variable targets investigates the time domain, and thus realizes the essence of a synoptic survey.  Variable and transient studies are all about sampling, and different types and time scales of phenomena benefit from different sampling strategies - sometimes significantly different.  Competing objectives described in this chapter are at the heart of LSST observing strategy and cadence design.

Target types are here grouped in subsections by variability characteristics, but as will be seen, this does not mean that all targets in a group require a common cadence, since the times scales may vary dramatically.  Acquiring suitable data for a wide range of time scales presents a fundamental problem for LSST, since the available $~$800 visits to a field over the survey cannot be deployed so as to usefully sample all time scales at all times.  This fact leads to the concept of a non-uniform survey, in which parts of the sky are visited more frequently part of the time.  The merits of such options must be traded against the benefits of a more uniform survey strategy.

When evaluating a particular observation or series of observations in light of how they perform for a specific science case, it may be helpful to think of metrics as lying along a continuum between discovery and characterization. Discovery requires a minimum amount of information to recognize an event or object as a candidate of interest, which necessarily involves some level of bare-bones characterization (upon which said recognition is based); rich characterization, on the other hand, implies that an event may not only be recognized as a candidate of interest, but basic properties of the event or object may be determined from the observation (e.g. including but not limited to classification of the event). The interpretation of a given metric along this continuum has implications for the subsequent action and analysis required, particularly as regards possible follow-up observations with other facilities. 


\section{Metrics for Transients and Variables}

As the science cases outlined below are grouped by temporal behavior (rather than underlying physical causes), the metrics developed to evaluate them may be generic and applicable to seemingly disparate science cases. Here, we describe metrics that are either already coded into the sims_maf_contrib repo or proposed. Individual subsections below should refer back to these metric descriptions.  


\subsection{Existing Metrics}

Lund et al. (2015; \url{http://arxiv.org/pdf/1508.03175.pdf}) discuss three metrics that have been incorporated into the MAF. Two of these metrics deal explicitly with time variable behavior: a) observational triplets, and b) detection of periodic variability. 

\subsubsection{Observational triplets (TripletMetric)}

This metric provides a means of evaluating whether a transient event on some timescale of interest has been detected, by testing for a sequence of three observations. The object must be detected in quiescence, followed by two subsequent detections above some threshold; this sequence of observations allows the magnitude of the change to be measured, as well as its timescale. 

This metric may be used for a variety of astrophysical phenomena, in particular transient events on variable objects (e.g. novae, stellar flares), in that it is general with respect to the amplitude of the brightness variation as well as the timescale of said change. The requirement of a detection prior to outburst does constrain it to objects that have already been detected in quiescence (in other words, not necessarily ``true'' transients), although there may be some cases where this is not the case (e.g. a supernova occurring on a previously detected galaxy). In practice, the time lapse between the first and second and second and third observations must be comparable (between 10$^2$ and 10$^5$ seconds) for discovery. This metric may be calculated for a given OpSim run and then further reduced to a histoogram in logarithmic time bins; the minimum number of bins to construct an interesting sample of objects is source-dependent. 

\subsubsection{Periodogram purity function (PeriodicMetric)}
This metric calculates the Fourier power spectral window function of each field (Roberts et al. 1987) as a means of quantifying the completeness of phase coverage for a given periodic variable. The periodogram purity is defined as 1 minus the Fourier power spectral window function; in the perfect case, all power is concentrated in a delta function at the correct frequency, and is zero elsewhere. As power ``leaks'' away from the correct frequency as a consequence of discrete, non-ideal data sampling, the periodogram becomes more structured. For the purposes of MAF metrics, which are designed to quantify performance as a single number, the periodogram purity is quantified as the minimum value away from the correct frequency. 

\subsubsection{Phase Gap Metric (PhaseGapMetric)}
Histogram of the median and maximum phase gaps achieved in all fields

\subsubsection{Period Deviation Metric (PeriodDeviationMetric)}

This metric computes the percent deviation of recovered periods from pure sine variability.

\subsubsection{Transient Metric (transientAsciiMetric)}

Calculate what fraction of transients would be detected using an ascii input file for the lightcurve.

\subsection{Proposed Metrics}

The following is a raw list of metric ideas; these need specificity and further description. 

The triplet metric may also be altered to include filter constraints, such that the triplets are drawn from a single filter or subset of filters.  

Color evolution constraint: triplets of observations in a specific color (really requirement of two triplets in multiple filters)

  2D Histogram of delta t?s between observations constituting a triplet 

FWHM of the window function (to quantify sampling)

Maximum hour angle difference 

Fraction of discoveries vs fractional duration of eclipse

Fraction of targets vs survey duration, for which the period can be determined to 5-sigma confidence

Fraction of targets vs survey duration, for which a period change of 1$\%$ can be determined with 5-sigma confidence

Histogram of median visit series length vs maximum visit spacing within the series

Number of events adequately sampled

% --------------------------------------------------------------------

\input{periodicvariablestars.tex}

% --------------------------------------------------------------------

\input{nonperiodicvariablestars.tex}

% --------------------------------------------------------------------

\input{periodictransients.tex}

% --------------------------------------------------------------------

\input{nonperiodictransients.tex}

% --------------------------------------------------------------------

% \input{cosmologicalSNe}

% --------------------------------------------------------------------

% \input{extragalactictransients}

% --------------------------------------------------------------------

\input{rollingcadence}

% --------------------------------------------------------------------
