\chapter[Synergy with WFIRST]{Synergy with WFIRST}
\def\chpname{wfirst}\label{chp:\chpname}

Chapter editor:
\credit{jasondrhodes}.

Contributing authors:
\credit{rubind},
\credit{davidpbennett},
\credit{mtpenny},
{\it Rachel Street}.

\section*{Summary}
\addcontentsline{toc}{section}{~~~~~~~~~Summary}

WFIRST will launch in $\sim$ 2025 for a 6 year mission to explore dark energy, find and characterize exoplanets, and take wide, deep infrared surveys of the galactic and extragalactic sky.  WFIRST was recognized by the Astro2010 Decadal Survey as an excellent NIR complement to LSST's optical capabilities. More recent work has recognized the strong synergy these two projects will have in helping to address cosmological questions in the 2020s \citep{2015arXiv150107897J}. 
  Together, the two observatories can accomplish significantly more (and better) science than either can alone in the areas of weak lensing, large-scale structure studies, strong lensing, supernova studies, exoplanet investigations,  and photometric redshift determination. Accomplishing this will require coordinated observations.  We have identified three areas of proposed coordination: 1. Early coverage of the $>2000$ square degree WFIRST High Latitude Survey to the full optical depth for enhanced photometric redshifts for both LSST and WFIRST; 2. Coordinated LSST optical observations in the WFIRST supernova discovery fields; 3. Precursor, simultaneous, and follow-up observations of the WFIRST microlensing fields near the galactic bulge.


% --------------------------------------------------------------------

\section{Introduction}
\label{sec:wfirst:intro}

% Introduce, with a very broad brush, this chapter's science projects,
% and why it makes sense for them to be considered together.

The Wide Field Infrared Survey Telescope (WFIRST) is a NASA mission that
entered Phase A in February 2016.  WFIRST was the highest recommendation
for large space missions in the 2010 New Worlds New Horizons Decadal
Survey.  That recommendation envisioned a wide-field observatory with
near infrared (NIR) capabilities to complement LSST's optical
capabilities; the Decadal Survey recognized the obvious synergy between
WFIRST and LSST.  WFIRST's design has evolved since 2010 and the design
being pursued for a mid-2020s launch uses an existing $2.4$m telescope
donated to NASA, giving WFIRST capabilities not envisioned by the
Decadal Survey.  WFIRST has 3 primary science objectives:

\begin{itemize}
\item Determine the nature of the dark energy that is driving the
current accelerating expansion of the universe using a combination of
weak lensing, galaxy clustering (including Baryon Acoustic Oscillations
and Redshift Space Distortions), and supernovae type Ia (SN).
\item Study exoplanets through a statistical microlensing survey and via
direct imaging and spectroscopy with a coronagraph.
\item Perform NIR surveys of the galactic and extragalactic sky via a
Guest Observer program.
\end{itemize}

WFIRST will be at L2 to enable the thermal stability needed for the
precise astrometric, photometric, and morphological measurements
required for these science goals. The baseline WFIRST mission
architecture is described in detail in the final report of the WFIRST
Science Definition Team (arxiv/1503.03757). The WFIRST Wide Field
Instrument(WFI) has a NIR focal plane with a $\sim0.28$ square degree
field of view made up 18 4k$\times$4k Teledyne H4RG NIR detectors and will
have imaging capabilities from $0.7-2$ microns and grism spectroscopy
capabilities from $1.35-1.89$ microns with $R\sim461\lambda$.  The WFI
also contains an Integral Field Channel (IFC) spectrometer with $R\sim100$
resoluton over the range $0.6-2$ microns for SN follow up. The exoplanet
coronagraph will have imaging ($0.43-0.97$ microns) and spectroscopic
($0.6-0.97 $ microns) capabilities with a contrast ratio of 1 part in a
billion.

WFIRST's  6 year primary mission is envisioned to have 2 years dedicated to a
$\sim2200$ square degree High Latitude Survey (HLS) for weak lensing and
galaxy clustering,  1 year of microlensing observations divided into 6
seasons, $0.6$ years of SN search and follow-up, one year dedicated to
the coronagraph and 1.4 years dedicated to competitively selected Guest
Observer observations. WFIRST has no expendables that would prevent an
extended mission of 10 years or longer, and an extended mission will likely be
given over entirely to Guest Observer observations.

The synergy with LSST is very promising indeed. In this chapter we aim
to  lay out three  specific projects in the three main WFIRST science
areas, and test the simulated LSST Observing
Strategies for their performance in each case. Then, we use these
results to design a suite of modified LSST Observing Strategies, which
we propose as new \OpSim simulation runs.



% --------------------------------------------------------------------

\input{WFIRST/WFIRST_weaklensing.tex}

% PJM: commented out pending check-in from Dave Rubin
\input{WFIRST/WFIRST_supernovae.tex}

\input{WFIRST/WFIRST_microlensing.tex}

% --------------------------------------------------------------------

\input{WFIRST/WFIRST_proposals.tex}

% --------------------------------------------------------------------

\navigationbar
