
% ====================================================================
%+
% SECTION:
%    periodicpulsators.tex
%
% CHAPTER:
%    variables.tex
%
% ELEVATOR PITCH:
%
%-
% ====================================================================

% \section{Discovery of Periodic Pulsating Variables}
\subsection{Discovery of Periodic Pulsating Variables}
\def\secname{periodicvariables}\label{sec:\secname}

\credit{lmwalkowicz},
\credit{StephenRidgway}

Regular variables, such as Cepheids and RR Lyraes, are valuable tracers
of Galactic structure and cosmic distance. In this case of these and
other strictly (or nearly-strictly) periodic variables, data from
different cycles of observation can be phase-folded to create a more
fully sampled lightcurve as LSST visits will occur effectively at random
phases. In a 10-year survey, most periodic stars of almost any period
will benefit from excellent phase coverage in all filters (only a very
small period range close to the sidereal day will be poorly observed).
Therefore, most implementations of the LSST observing strategy will
provide good sampling of periodic variables.

However, different implementations of the survey may result in different
resulting sample sizes of these periodic variables, and may also affect
the environments in which these stars are discovered. In this section,
we create a framework for understanding how current implementations of
the observing strategy influence (or even bias) the resultant sample
size and environments where these important tracers may be identified.

\subsubsection{Tracing Galactic Structure with RR Lyrae}

RR Lyrae variables are crucial tracers of structure in the Galaxy and beyond
into the Local Group. The incredible sample of RR Lyrae anticipated from LSST
observations will enable discovery of Galactic tidal stream and neighboring
dwarf galaxies throughout much of the Local Group
\citep{IvezicEtal2008}.  LSST also creates the possibility
of detecting and studying RRL variables in the Magellenic Clouds; see Chapter
[CHAPTER] for discussion.

\citet{2012AJ....144....9O} carried out an extensive simulation of period
and lightcurve shape recovery of RR Lyrae variables using an early OpSim run
opsim1$\_$29. Correctly identifying the period aids in building the sample of
interest, whereas fitting the lightcurve shape makes it possible to measure the
metallicity of the star. In their simulation, they employed both a Fourier
analysis and template matching to recover the lightcurve shape, finding that
template matching yielded a more accurate lightcurve shape measurement in the
presence of sparse data. The results of this simulation showed that the vast
majority of RR Lyrae will be discovered by the baseline observing strategy (as
deployed in opsim1$\_$29) within 5 years of survey operations. Half of both
RRLab and RRLc stars will be found out to $\sim$600 kpc and $\sim$250 kpc
(respectively) by the end of the 10-year main survey, and template matching
techniques for lightcurve shape recovery will provide metallicities to
$\sim$0.15dex.

% --------------------------------------------------------------------

\subsubsection{The Cepheid Cosmic Distance Ladder}

Classical cepheids remain an essential step in the cosmic distance
ladder. Their calibration is based largely on LMC cepheids and known
(assumed) distance of the LMC.  The associated errors, while uncertain,
are believed to be of $\>\sim$7\%. (Madore, Barry F.; Freedman, Wendy L.
(2009). "Concerning the Slope of the Cepheid Period?Luminosity
Relation". The Astrophysical Journal 696 (2): 1498. arXiv:0902.3747.
Bibcode:2009ApJ...696.1498M. doi:10.1088/0004-637X/696/2/1498.) New
developments in galactic studies are poised to support substantially
improved descriptive information concerning nearby galactic cepheids,
with possible substantial reductions in this error, by accurately
securing the PL slope and zero point.

Cepheid calibration errors are associated in part with uncertainties in
extinction, both interstellar and in some cases circumstellar, and in
metalicity.  At present, the direct, local calibration of cepheids is limited
by the availability of a few direct distance measurements, obtained with HST,
with errors $\sim$10\%.  The GAIA mission is expected to return $\sim$9000
Galactic cepheids, of all periods, colors and metallicities, with distance
errors less than 10\% (many of them much less) - Windmark, F.; Lindegren, L.;
Hobbs, D., 2011A\&A...530A..76W. It is expected to deliver at least 1000
cepheids in the LMC with expected mean distance error $\sim$7-8\% (Clementini
(2010) - 011EAS....45..267C).  GAIA, as well as other methods, will also
support determination of the 3-d map of galactic interstellar extinction -
including possible variations in the extinction law. These rich data sets will
be supported with direct measurements of cepheid diameters (A. Merand et al,
A\&A in press) and advances in stellar hydrodynamics
\citep{2013MNRAS.435.3191M} which will provide theoretical and empirical basis
for calibrations to reconcile known physics with observational correction
factors.

Galactic cepheids will generally be too bright for LSST, but cepheids in
the local group are sufficiently bright that LSST photometry will be
limited by calibration errors rather than by brightness.  This dataset
will provide superb support for integration of GAIA-based galactic
cepheid studies with extra-galactic cepheid studies.

GAIA will provide similar precision data with the potential to identify
or support distance determinations from many other galactic star types.
LSST photometric catalogs will represent a uniquely extensive and
complete database for such investigations.

% --------------------------------------------------------------------

% \subsection{Metric Analysis}
\subsubsection{Metric Analysis}
\label{sec:\secname:analysis}

Several metrics currently exist in the MAF for evaluating how LSST
survey strategy affect the recovery of periodic sources.
For example,
the PeriodicMetric makes use of the periodogram purity function, which effectively
quantifies aliasing introduced into periodogram analysis from the
sampling of the lightcurve.
%
% and period deviation metric (PeriodDeviationMetric) all return relevant
% information
%
Similarly, the
phase gap metric (PhaseGapMetric),
evaluates the periodicity of the source lightcurve and its coverage
in phase space (the latter being relevant for shape recovery).

Recreating the template matching results of the
\citet{2012AJ....144....9O} simulation requires sampling specific input
lightcurves and comparing with the library of available shapes; this
necessarily requires a step outside of the MAF, but can easily be
enabled using the lightcurve simulation tools under development.
%  [NAME OF FED'S LIGHTCURVE
% TOOL].

Current simulations of the main survey show a broad uniformity of
visits, with thorough randomization of visit phase per period, giving
very good phase coverage with minimum phase gaps.


% % --------------------------------------------------------------------
%
% \subsection{Discussion}
% \label{sec:\secname:discussion}

For periodic variable science, two cadence characteristics should be avoided:
\begin{itemize}
\item an exactly uniform spacing of visits (which is anyway virtually impossible); \
\item a very non-uniform distribution, such as most visits concentrated in a few survey years.
 \end{itemize}

A metric for maximum phase gap will guard against the possibility that a
very unusual cadence might compromise the random sampling of periodic
variables.

In each case, it would help to jump-start science programs if some
fraction of targets had more complete measurements early in the survey.

% ====================================================================
%
% \subsection{Conclusions}
%
% Here we answer the ten questions posed in
% \autoref{sec:intro:evaluation:caseConclusions}:
%
% \begin{description}
%
% \item[Q1:] {\it Does the science case place any constraints on the
% tradeoff between the sky coverage and coadded depth? For example, should
% the sky coverage be maximized (to $\sim$30,000 deg$^2$, as e.g., in
% Pan-STARRS) or the number of detected galaxies (the current baseline but
% with 18,000 deg$^2$)?}
%
% \item[A1:] No strong constraint - but see Q4.
%
% \item[Q2:] {\it Does the science case place any constraints on the
% tradeoff between uniformity of sampling and frequency of  sampling? For
% example, a rolling cadence can provide enhanced sample rates over a part
% of the survey or the entire survey for a designated time at the cost of
% reduced sample rate the rest of the time (while maintaining the nominal
% total visit counts).}
%
% \item[A2:] An enhanced cadence could provide earlier discovery and period confirmation for a subset of targets, but this is
%not a high priority.
%
% \item[Q3:] {\it Does the science case place any constraints on the
% tradeoff between the single-visit depth and the number of visits
% (especially in the $u$-band where longer exposures would minimize the
% impact of the readout noise)?}
%
% \item[A3:] No strong constraint.
%
% \item[Q4:] {\it Does the science case place any constraints on the
% Galactic plane coverage (spatial coverage, temporal sampling, visits per
% band)?}
%
% \item[A4:] Most stars are in the galactic plane. For periodic pulsators, it is desirable to obtain multiple visits
%in all filters, with at least 20-50 epochs in several filters. Obtaining this coverage is more important to periodic 
%pulsator science than greater sky coverage.
%
% \item[Q5:] {\it Does the science case place any constraints on the
% fraction of observing time allocated to each band?}
%
% \item[A5:] No strong constraint, as long as there is a good representation of each, including u.
%
% \item[Q6:] {\it Does the science case place any constraints on the
% cadence for deep drilling fields?}
%
% \item[A6:] For deep drilling on the galaxy or nearby galaxies, the best cadences would cover a range of periods from
%minutes to days - longer periods would be satisfactorily sampled by the main survey.
%
% \item[Q7:] {\it Assuming two visits per night, would the science case
% benefit if they are obtained in the same band or not?}
%
% \item[A7:] Different bands would provide more rapid characterization of targets, but this is not a strong benefit.
%
% \item[Q8:] {\it Will the case science benefit from a special cadence
% prescription during commissioning or early in the survey, such as:
% acquiring a full 10-year count of visits for a small area (either in all
% the bands or in a  selected set); a greatly enhanced cadence for a small
% area?}
%
% \item[A8:] Enhanced cadences could provide earlier science, or somewhat deeper science, depending on details, but 
%periodic variables are not a strong driver for this.
%
% \item[Q9:] {\it Does the science case place any constraints on the
% sampling of observing conditions (e.g., seeing, dark sky, airmass),
% possibly as a function of band, etc.?}
%
% \item[A9:] No constraints that are particular to variable stars.
%
% \item[Q10:] {\it Does the case have science drivers that would require
% real-time exposure time optimization to obtain nearly constant
% single-visit limiting depth?}
%
% \item[A10:] No.
%
% \end{description}

% ====================================================================

\navigationbar
